### 2023-04-03 manuscript/manuscript.tex
manuscript LaTeX backmatter, packages, commands

\documentclass[11pt, 5p]{elsarticle}

\journal{TBA}

\bibliographystyle{elsarticle-num}
\biboptions{numbers,sort&compress}

% format hacks
\usepackage{libertine}
\usepackage{libertinust1math}
\usepackage{amsmath}                 % Provides advanced math typesetting and environments
\usepackage{bbold}                   % Provides blackboard bold font for numbers
\usepackage{graphicx}                % Handles the inclusion of graphics files
\usepackage{eurosym}                 % Provides the Euro currency symbol
\usepackage{mathtools}               % Enhances amsmath package, adding new symbols and tools
\usepackage{url}                     % Handles the formatting of URLs
\usepackage{booktabs}                % Improves the quality of tables in LaTeX
\usepackage{epstopdf}                % Automatically converts EPS files to PDF
\usepackage{xfrac}                   % Provides a consistent way to typeset fractions
\usepackage{tabularx}                % Enhances the standard LaTeX tables
\usepackage{bm}                      % Provides access to bold math symbols
\usepackage{subcaption}              % Provides support for subfigures and subtables
\usepackage{longtable}               % Allows the creation of multi-page tables
\usepackage{multirow}                % Enables cells spanning multiple rows in tables
\usepackage{threeparttable}          % Provides additional functionality for tables
\usepackage{pdflscape}               % Allows landscape pages in PDF output
\usepackage[svgnames]{xcolor}    	 % Load xcolor packages
\usepackage[export]{adjustbox}       % Provides tools for resizing and trimming images
\usepackage[version=4]{mhchem}       % Provides typesetting for chemical equations
\usepackage[colorlinks]{hyperref}    % Creates hyperlinks in the document
\usepackage[parfill]{parskip}        % Adjusts paragraph spacing
\usepackage[nameinlink,sort&compress,capitalise,noabbrev]{cleveref} % Enhanced cross-referencing
\usepackage[leftcaption,raggedright]{sidecap} % Provides side captions for figures and tables
\usepackage[prependcaption,textsize=footnotesize]{todonotes} % Allows inline and margin notes
\usepackage{siunitx}                 % Provides consistent formatting for units and numbers
\usepackage[acronym, automake, nonumberlist]{glossaries-extra}

\usepackage{lipsum}
\usepackage{mdframed}
\usepackage[resetlabels,labeled]{multibib} % Allows multiple bibliographies with own styles/numbering
\newcites{S}{Supplementary References}
\bibliographystyleS{elsarticle-num}

\sisetup{
	range-units = single,
	per-mode = symbol
}
\DeclareSIUnit\year{a}
\DeclareSIUnit{\tco}{t_{\ce{CO2}}}
\DeclareSIUnit{\sieuro}{\mbox{\euro}}

\graphicspath{{../results/graphics/},}

\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\set}[1]{\left\{#1\right\} }
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newcommand{\R}{\mathbb{R} }
\newcommand{\B}{\mathbb{B} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\co}{\ce{CO2}~}
\def\el{${}_{\textrm{el}}$}
\def\th{${}_{\textrm{th}}$}
\def\deg{${}^\circ$}

\newdefinition{rmk}{Remark}
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newmdtheoremenv[
  backgroundcolor=Wheat, % Background color
  linewidth=0.5pt, % Width of the line around the box
  frametitlefont=\normalfont\bfseries, % Font for the box title
]{res}[thm]{Result}

% acronyms
\makeglossaries
\newacronym{tes}{TES}{Thermal energy storage (in form of water tanks)}
\newacronym{chp}{CHP}{Combined heat and power plant}
\newacronym{cop}{COP}{Coefficient of Performance}
\newacronym{bdew}{BDEW}{Bundesverband der Energie- und Wasserwirtschaft}
\newacronym{PV}{PV}{Solar photovoltaics}
\newacronym{ghg}{GHG}{Greenhouse gas}
\newacronym{EU}{EU}{European Union}
\newacronym{dea}{DEA}{Danish Energy Agency}
\newacronym{helmeth}{HELMETH}{Integrated High-Temperature Electrolysis and Methanation for Effective Power to Gas Conversion}
\newacronym{jrc}{JRC}{Joint Research Center}
\newacronym{ocgt}{OCGT}{Open cyclic gas turbine}
\newacronym{v2g}{V2G}{Vehicle to Grid}
\newacronym{pypsa}{PyPSA}{Python for Power System Analysis}
\newacronym{tyndp}{TYNDP}{Ten-year network development plan}
\newacronym{entso-e}{ENTSO-E}{European network for transmission system operators electricity}
\newacronym{nuts}{NUTS}{Nomenclature of Territorial Units for Statistics}
\newacronym{covid}{COVID-19}{Coronavirus disease 2019}
\newacronym{entsog}{ENTSO-G}{European Network of Transmission System Operators for Gas}
\newacronym{dsm}{DSM}{Demand side management}
\newacronym{hvdc}{HVDC}{High voltage direct current}
\newacronym{phs}{PHS}{Pumped hydro storage}
\newacronym{dac}{DAC}{Direct air capture}
\newacronym{sng}{SNG}{Synthetic natural gas}
\newacronym{fom}{FOM}{Fixed operation and maintenance costs}
\newacronym{vom}{VOM}{Variable operation and maintenance costs}
\newacronym{ppa}{PPA}{Power Purchase Agreement}
\newacronym{ptc}{PTC}{Production Tax Credit}
\newacronym{res}{RES}{Renewable energy sources}
\newacronym{necp}{NECP}{National energy climate plan}
\newacronym{ets}{ETS}{Emissions Trading System}
\newacronym{capex}{CAPEX}{Capital expenditures}
\newacronym{cfd}{CfD}{Contract for Difference}
\newacronym{cfe}{CFE}{Carbon-free electricity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

	\title{ new paper}

	\author[affil]{ iegor } % \corref{correspondingauthor}
	\ead{ name@box }

	\address[affil]{ affiliation }

	\begin{abstract}
		\input{sections/abstract.tex}
	\end{abstract}

	\begin{keyword}
		TODO
	\end{keyword}

	% \begin{graphicalabstract}
	% \end{graphicalabstract}

\end{frontmatter}

%\listoftodos[TODOs]
%\tableofcontents

\section{Introduction}
\label{sec:intro}
\input{sections/introduction.tex}

\section{Methods}
\label{sec:methods}

\input{sections/methods.tex}

\section{Results}
\label{sec:results}

\input{sections/results.tex}

\section{Discussion}
\label{sec:discussion}

\input{sections/discussion.tex}

\section{Conclusion}
\label{sec:conclusion}

\input{sections/conclusion.tex}

\section*{Acknowledgements}

% XY gratefully acknowledge funding from XY under grant no. 00000.

\section*{License}
% \doclicenseLongText
% \doclicenseIcon

\section*{Author Contributions}

% following https://casrai.org/credit/

\textbf{Author XY}:
Conceptualization --
Data curation --
Formal Analysis --
Funding acquisition --
Investigation --
Methodology --
Project administration --
Resources --
Software --
Supervision --
Validation --
Visualization --
Writing - original draft --
Writing - review \& editing

\section*{Data and Code Availability}

A dataset of the model results is available at \href{zenodoTBA}{doi:zenodoTBA}.
The code to reproduce the experiments is available at \href{https://github.com/repository}{GITHUB}.

\section*{Abbreviations}
\printglossary[type=\acronymtype]

% tidy with https://flamingtempura.github.io/bibtex-tidy/
\addcontentsline{toc}{section}{References}
\renewcommand{\ttdefault}{\sfdefault}
\bibliography{references}

% supplementary information

\newpage

\makeatletter
\renewcommand \thesection{S\@arabic\c@section}
\renewcommand\thetable{S\@arabic\c@table}
\renewcommand \thefigure{S\@arabic\c@figure}
\makeatother
\renewcommand{\citenumfont}[1]{S#1}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}

\section{Supplementary Information}
\label{sec:si}

\input{sections/supplementary.tex}

\addcontentsline{toc}{section}{Supplementary References}
\renewcommand{\ttdefault}{\sfdefault}
\bibliographyS{references}

\end{document}