### 2023-04-04 manuscript/sections/abstract.tex

Traditional power purchase agreements for renewable energy have seen rapid growth in recent years, but they only match supply and demand on average over a longer period such as a year.
There is increasing interest from leaders in voluntary clean electricity procurement to cover their consumption with clean energy supply on truly 24/7 basis. 
Achieving 24/7 carbon-free energy means that every kilowatt-hour of electricity consumption is met with carbon-free electricity sources, every hour of every day.
In this study, we investigate both the means and costs of pursuing different clean electricity procurement strategies for companies in a selection of European countries.
We also explore how the 24/7 clean energy procurement affects the rest of the European electricity system.

\bf{Dummy abstract - to refine}