### 2023-04-04 manuscript/sections/conclusion.tex

Taken together, we show that 24/7-CFE commitment by corporate sector stakeholders is a step above and beyond the 100\% annual renewable matching in paving the path toward the carbon net-zero energy systems. 
This voluntary commitment also delivers a range of benefits to the society by 
(i) triggering innovation, availability and earlier utilisation of advanced clean firm generation needed for the transformation of energy systems to a complete carbon net-zero state, 
(ii) reducing system emissions, and (iii) reducing system flexibility needs.