### 2023-10-10 manuscript/sections/supplementary_2.tex

%S2: Background system assumptions
In all the modelled countries, renewable generation must meet the political targets as defined in the NECPs or by more recent national policy targets (such as the Easter package in Germany), see Table 1.