### 2023-10-10 manuscript/sections/supplementary_3.tex

%S3: Mathematical formulation of clean energy procurement
\cite{srivastavCarbonCaptureUtilisation} %test supplementary references