### 2023-10-14 manuscript/sections/conclusion.tex
manuscript: add conclusions

24/7 carbon-free energy matching is a viable and effective strategy for electricity buyers aiming to eliminate their own carbon footprint and contribute to wider system decarbonisation.
Voluntary commitments to 24/7 CFE procurement have further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.
The evidence holds across geographical locations and time periods, and is robust to sensitivity analyses.
The strategy retains its significance in reducing system-level emissions, even as grids become cleaner over time, which is essential in light of evolving national energy and climate policies.