### 2023-10-15 manuscript/sections/abstract.tex
manuscript: add abstract

% The 5 things a good abstract needs:

% 1. Introduce the topic: what we know 
% 2. State the unknown: what we don't know
% 3. Outline the methods: how we answer the questions
% 4. Preview the findings: what we learned
% 5. Tell us what your work teaches us: what we conclude and why this is important

% aim ~200 words | 225 words used

A growing number of public and private energy buyers are interested in 24/7 carbon-free energy (CFE) procurement, which means that every kilowatt-hour of electricity consumption is met by carbon-free sources at all times.
It has the potential to overcome the limitations of established procurement schemes, such as the temporal mismatch between clean electricity supply and buyers' demand that is inherent to "volumetric" matching. 
Yet it is unclear whether 24/7 CFE procurement is only effective during the early stages of the energy transition, or if it has a system-level impact in a variety of regional contexts.
We use a mathematical model to systematically examine different designs, optimal procurement strategies, costs, and impacts of the 24/7 CFE matching, both for participating buyers and for regions where voluntary procurement occurs.
We examine the mechanisms driving system-level co emissions reduction and how they vary across regions and over time.
Our results indicate that voluntary clean energy procurement commitments have consistent effects on participants and the electricity system.
Even as grids become cleaner over time, the hourly matching strategy remains significant in reducing CO$_2$ emissions.
In addition, voluntary commitments to 24/7 CFE have a further transformative effect on electricity systems through accelerated innovation and early deployment of advanced energy technologies.
This work generalises the implications of clean energy procurement strategies in the context of evolving national energy and climate policies.