### 2024-07-17 manuscript/sections/conclusion.tex

% SPDX-FileCopyrightText: 2023 Iegor Riepin, Tom Brown
%
% SPDX-License-Identifier: CC-BY-4.0

In this work, we systematically examined the means, costs, and the system-level impacts of the 24/7 carbon-free energy matching.
We find that 24/7 matching is a viable and effective strategy for clean electricity buyers aiming to eliminate their own carbon footprint and contribute to wider system decarbonisation.
The evidence holds across different geographical locations.
Even as grids become cleaner over time, the 24/7 matching strategy retains its significance in reducing system-level emissions, which is essential in light of evolving national energy and climate policies.
Voluntary commitments to the hourly matching strategy have a further transformative effect on electricity systems by creating an early market for advanced energy technologies, such as long-duration energy storage and clean firm power generators.